### 2025-01-15 paper/rebuttal.txt
Initial draft of rebuttal
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

First, we thank all the reviewers for their time, their careful reviews,
and their thoughtful feedback.

Several reviewers suggested citations that we are happy to add.

Other reviewers asked how we handle various litmus tests.  Our submission
describes how we rely on a compiler to translate the litmus test to object
code, then relate dependencies in that object code to the corresponding
source code.  Section 1.2 ("Prior Work") notes that we are not trying
to identify when OOTA cycles can form, but rather that OOTA cycles
cannot form.

Several reviewers seemed uncomfortable with informal proofs.  However,
ASPLOS has published papers with informal proofs, so we suspect that the
real discomfort was augmenting the C++ abstract machine with hardware
properties.  20 years of equivocal progress is more than sufficient to
justify trying a new approach.  We are not advocating the old approaches
be abandoned.  In fact, should someone succeed with those approaches,
we will endeavor to be the first to offer congratulations.

Others were uncomfortable with our reliance on object code, which
varies across systems.  However, these systems are underpinned by laws
of physics, namely causality, the finite speed of light, and the atomic
nature of matter.  Our definition of OOTA cycles as cycles in (rfe U
sdep) permits us to examine rfe and sdep separately.  These laws of
physics guarantee that each rfe link goes forward in time.  Because an
OOTA cycle must return to the starting time, there must be at least
one sdep link that goes backwards in time.  Correct compilers guarantee
that sdep links correspond to sequences of machine instructions, which
cannot execute backwards in time, even given speculative execution.
Thus, the OOTA cycle cannot form.

Please note that correct implementations are not required to break
non-semantic dependencies.  An unbroken non-semantic dependency will
correspond to a sequence of machine instructions that cannot go backwards
in time, and thus cannot form an OOTA cycle.

Reviewer 72F asks about three litmus tests in Section 5 of the
submission's citation [5], as do a couple of other reviewers:

o	The first example has a dependency only if the implementation
	does not hoist the i++ out of the “if” statement, and this
	is the implementation’s choice, similar to the example in
	Section 2.2.5 of the submission.

o	The second example is handled by examining f() and the functions
	that it invokes, along with the implementation’s optimization
	choices.  Note that at one extreme, f() might ignore its
	parameter and at the other extreme it might simply return
	its parameter.  Note further that f() might be a pointer to a
	function, and that function might come from a dynamically linked
	library that has not yet been written at the time the call to
	f() is compiled.  Although the submission does not contain
	this particular example, all of these issues are addressed by
	the submission’s Section 3, which discusses executions (as
	opposed to source-code syntactic dependencies).

o	An instructive elaboration of the third example is discussed in
	Section 2.2.1 of the submission.

Reviewer 72F also asks if our approach implicitly enforces the
load-to-store ordering advocated by Boehm and Demsky.  The answer is
"no".  First, although instructions commit in some order, they do not
necessarily commit in program order, which on some systems allows prior
loads to be reordered with later stores.  Even on hardware that does
commit instructions in program order, there is no guarantee that this
commit order will constrain the visibility of the corresponding memory
references, for example, due to the effects of hardware store buffers.

Reviewer 72F also asks (in two questions) if freezing of nondeterministic
choices into machine code is realistic.  The answer is "yes", but on the
per-execution basis used in our submission.  For example, specialization
optimizations might cause a given function to be compiled differently
for different combinations of arguments.  But any given execution would
correspond to a particular combination of arguments, and that execution
would be frozen into machine code.  Please note that programs that
modify their own code are outside of the standard, and thus out of our
submission's scope.

We especially thank Reviewer 72F for noting (correctly!) that Section 7.4
("Merging Quasi-Volatile Loads") poses a challenge, as this helped refine
our thinking. The refinement is to note that this challenge is not to our
technique, but rather to those implementing optimizations.  Although we
believe (one of us strongly, the rest less so) that the "optimization"
described in this section is in fact a pessimization that also degrades
ease of use, the question of whether or not this optmization is legitimate
will be decided by community processes, not by fiat and not by pure
mathematics.  Nevertheless, should an abstract-machine mathematical
treatment of OOTA appear, we would expect that this mathematics would
have a prominent seat at the community-process table.  Either way,
if this optimization is rejected the litmus test in Section 7.4 has a
semantic dependency, otherwise it does not.  And correct implementations
will reflect that decision, though again, optimizations are permitted
to refrain from breaking non-semantic dependencies.

Reviewer 72F suggested that rf links must impose restrictions.
Please note our Section 2.1 ("OOTA: rf vs. rfe") where we show that
those restrictions can be extremely loose, hence our use of rfe.
But note also that C++ modification order provides very weak guarantees,
in fact, a later-in-time store might precede an earlier-in-time store
in modification order, even on x86.  Memory models are at best loosely
connected to the passage of time.

Reviwer 72F also questions how the original C++ code can be irrelevant.
The C++ code is relevant, but only at the source-code level and to the
connections between the source-code and object-code levels.  However,
for analysis confined to the object-code level, the origial C++ code
can indeed be irrelevant.

Reviewer 72F asks what we mean by "these models are atemporal".  We
mean that typical memory models do not make use of the passage of
physical time.  The provide partial orderings, but as noted earlier,
the modification order can in fact go backwards in time.

Responses to individual reviewer feedback items is in the accompanying
PDF.