### 2025-01-15 paper/rebuttal.txt
rebuttal.txt:  Apply Alan Stern feedback
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

First, we thank all the reviewers for their careful reviews and thoughtful
feedback.

Responses to individual reviewer feedback items may be found in the
accompanying PDF.

We are happy to add the suggested citations.

There was discomfort with informal proofs.  However, our proofs are
straightforward and ASPLOS has published such proofs.  We therefore
suspect that the real discomfort was augmenting the C++ abstract machine
with hardware properties.  20 years of equivocal progress justifies trying
a new approach.  We are not advocating the old approaches be abandoned.
In fact, we will endeavor to be the first to offer congratulations should
someone succeed.

Various litmus tests were put forward.  Our submission describes how
we rely on a compiler to translate the litmus test to object code, then
relate dependencies in that object code to the corresponding source code.
Section 1.2 ("Prior Work") notes that we are not trying to identify when
OOTA cycles can form, but rather show that OOTA cycles cannot form.

There was also discomfort with use of object code, which varies across
systems.  However, these systems are underpinned by laws of physics,
namely causality, the finite speed of light, and the atomic nature
of matter.  Our definition of OOTA cycles as cycles in (rfe U sdep)
permits us to examine rfe and sdep separately, noting that each rfe link
goes forward in time.  Because an OOTA cycle must return to the starting
time, there must be at least one sdep link that goes backwards in time.
Correct compilers guarantee that sdep links correspond to sequences of
machine instructions, which cannot execute backwards in time, even given
speculative execution.  Thus, the OOTA cycle cannot form.

Please note that correct implementations are not required to break
non-semantic dependencies.  Regardless, a non-semantic dependency cannot
form an OOTA cycle.

Responses specifically requested by Reviewer 72F for the rebuttal:

How do we handle the three litmus tests in Section 5 of citation [5]?

o	The first example has a dependency only if the implementation
	does not hoist the i++ out of the “if” statement, and this is
	the implementation’s choice, similar to the example in Section
	2.2.5 of the submission.  See Sections 2.2.7 and 2.2.8 for more
	aggressive examples.

o	The second example is handled by examining f() and the
	functions that it invokes, along with the implementation’s
	optimization choices.  This example is yet another motivation
	for the submission’s Section 3, which discusses executions as
	opposed to source-code syntactic dependencies.

o	An instructive elaboration of the third example is discussed in
	Section 2.2.1 of the submission.

Do we implicitly enforce the load-to-store ordering advocated by Boehm
and Demsky?  (Two questions.)  We do not.  Instructions do not necessarily
commit in program order.  Furthermore, commit order might not constrain
the order of the corresponding memory references.

Is freezing of nondeterministic choices into machine code realistic?
Yes, but on the per-execution basis used in our submission.  For example,
specialization optimizations might cause a given function to be compiled
differently for different combinations of arguments.  But any given
execution would correspond to a particular combination of arguments,
and that execution would be frozen into machine code.

We thank Reviewer 72F for noting that Section 7.4 ("Merging Quasi-Volatile
Loads") poses a challenge.  However, please note that this challenge
is not to our technique, due to our focus on object-code executions.
Community processes will decide whether or not this section's optimization
is legitimate, and these processes' outcomes will be implemented in
compilers, thus being automatically handled by our technique.  The
challenge is instead to OOTA analysis techniques not accounting for these
community processes.  Nevertheless, any abstract-machine mathematical
OOTA solution should have a prominent seat at the community-process table.

Don't rf links must impose restrictions?  Yes, they do in the C++ memory
mode, but only when they involve release-stores and acquire-loads or
consume-loads.  These restrictions do not apply to OOTA cycles, which
involve only relaxed accesses.  Note further that C++ modification
order provides very weak guarantees, in fact, a later-in-time store
might precede an earlier-in-time store in modification order, even on x86.

How the original C++ code can be irrelevant?  The C++ code is relevant,
but only at the source-code level and to the connections between the
source-code and object-code levels.  However, for analysis confined to
the object-code level, the original C++ code can indeed be irrelevant.
For example, if a program in C++ and another in Rust generated the same
machine code, the behavior of that machine code would be independent of
the source language.

What is meant by "these models are atemporal"?  We mean that typical
memory models do not make use of the passage of physical time.
The provide partial orderings, but as noted earlier, some orderings can
go backwards in time.