### 2025-01-16 paper/rebuttal-mods.txt
rebuttal-mods.txt: Track possible upcoming changes
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

The "o" items assume a venue (such as a tech report) that have relaxed
page limits.

O	Detailed feedback from Review #72C.

O	s/We would be happy to change "explicitly prohibits"/"explicitly recommends against"/
	This is prior to the quote from the C++ standard.

O	s/Workers/Researcher/ in Section 1.2.

O	Add citations to Section 1.2 ("Prior Work").

O	Be less condescending towards Mark Batty.  Maybe
	s/is persisting with/is working on/.

o	Add a Section 1.3 ("Our Approach"?) that summarizes the
	rfe/sdep/non-sdep-dep approach and emphasizes proving
	OOTA cannot happen as opposed to an exact delineation
	between sdep and non-sdep.

o	Add citations to later sections that discuss similar litmus tests.

o	Consider adding a 2.2.x section that discusses the difficulties
	inherent in computing sdep for function calls, especially when
	the function will be part of a dynamic library that has not yet
	been written.

O	In section 6.1, define "witness" informally as "another execution
	in which the store acts differently".

O	On page 9, "While it possible".  [ Since fixed by Alan. ]

O	In Section 7.4 ("Merging Quasi-Volatile Loads"), change focus
	from intuitive notions to a community-process approach for
	determining what optimizations are legitimate.  Math cannot
	handle this.  We avoid the problem by looking at the compiler
	implementation, which is downstream of the community process.

	Also in Section 7.4 change "our intuitive notions" to "one's
	intuitive notions".

o	Possible future work: Relax quasi-volatile semantics to permit
	additional uses of the cmov instruction.